%-------------------------------------------------------------------------------
% PREAMBLE AND DOCUMENT FORMATTING
%-------------------------------------------------------------------------------
\documentclass[journal=ancac3, manuscript=suppinfo, etalmode=truncate,maxauthors=0]{achemso}
\setkeys{acs}{etalmode=truncate,maxauthors=0}

% PACKAGES
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{textgreek}

\usepackage[version=4]{mhchem}
\usepackage[alsoload=synchem,separate-uncertainty=true,multi-part-units=single,tight-spacing=true]{siunitx}

\usepackage{float}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage{array, booktabs, tabularx, multirow}

% CAPTION FORMATTING
\usepackage[font=footnotesize,labelfont=bf,labelsep=period]{caption} 							% format single-image captions and table titles
\captionsetup[table]{singlelinecheck=false,font=footnotesize,labelfont=bf}
\usepackage[font=footnotesize,labelfont=bf,labelsep=period]{subcaption} 						% format subfigure captions
%\DeclareCaptionSubType*[alph]{figure}
%\renewcommand\thesubfigure{\thefigure\alph{subfigure}}
\captionsetup[subfigure]{labelfont=bf,textfont=normalfont,labelformat=simple,singlelinecheck=false} 	

% CROSS-REFERENCE FORMATTING
% For use with the cleveref package
% Define the format of Figure, Table, Equation, and Section cross-references in the text
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{cleveref}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Tab.}{Tabs.}
\Crefname{table}{Table}{Tables}
\crefname{equation}{Eq.}{Eqs.}
\Crefname{equation}{Equation}{Equations}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}


% REFERENCES
\renewcommand*{\bibfont}{\normalfont\small}


%-------------------------------------------------------------------------------
% CUSTOM COMMANDS
%-------------------------------------------------------------------------------


\input{additional_commands.tex}


%-------------------------------------------------------------------------------
% TITLE AND AUTHOR LIST
%-------------------------------------------------------------------------------
\title{Modelling of Ion and Water Transport in the Biological Nanopore ClyA}
\input{authors.tex}


%===============================================================================
%
% MAIN DOCUMENT BEGINS
%
%===============================================================================
\begin{document}
	
\maketitle

\newpage
\section{Extended materials and methods}

\subsection{ClyA-AS homology model}
A full atom model of ClyA-AS\cite{Soskine-2013} was build using the MODELLER (version 9.18) software package 
by introduction of the following point mutations in each of the 12 chains of the wild-type ClyA crystal 
structure (PDBID: 
2WCD\cite{Mueller-2009}):
K8Q, N15S, Q38K, A57G, T67V, C87A, A90V, A95S, L99Q, E103G, K118R, L119I, I124V, T125K, V136T, F166Y, K172R, 
V185I, K212N, K214R, S217T, T224S, N227A, T244A, E276G, C285S, K290Q.
The resulting structure was optimized by annealing first the mutated residues alone, and subsequently also 
the closed atoms around them, using a conjugate gradient (from $150$ to $1000$ to $300$~K with $4$~fs 
timesteps).\cite{Sali-1993}

% are listed in \cref{tab:clya_as_mutations}.
% \input{tab_clya_as_mutations.tex}

\newpage
\subsection{Fitting of the electrolyte properties}
\input{fig_corrections.tex}
The parameters of the fitting functions used to interpolate the experimental ion diffusion coefficients, 
mobilities and transport numbers, and the electrolyte viscosity, density and relative permittivity are given 
in \cref{tab:corrections_parameters} and the resulting curves are plotted in \cref{fig:corrections}. Note 
that since most of these functions are merely empirical fits with no physical meaning, and that they are 
solely used to interpolate and represent the experimental data.

Finally, for the concentration dependence of the relative permittivity we made use of the model proposed by Gavish et al.:
\begin{align}
\relperm(\dconc) = \permittivity_{r,0} - \left(\permittivity_{r,0} - \permittivity_{r,ms}\right) L \left( \dfrac{3\alpha}{\permittivity_{r,0} - \permittivity_{r,ms}} \dconc \right)
\end{align}
\input{tab_corrections_parameters.tex}

\newpage
\subsection{Surface integration to compute pore averaged values}
The average pore values for quantity of interest $X$ was computed by
\begin{align}
  \left< X \right>_{\alpha} =
    \displaystyle\frac{\displaystyle\iint_{V_{\alpha}} \beta_{\alpha} X \,dr\,dz}
                      {\displaystyle\iint_{V_{\alpha}} \beta_{\alpha} \,dr\,dz}
  \text{ ,}
\end{align}
where
\begin{equation}
  \alpha=
  \begin{cases}
    \text{p}, & d \ge 0  \text{~nm} \text{, average over the entire pore} \\
    \text{b}, & d > 0.5  \text{~nm} \text{, average over the pore `bulk' }  \\
    \text{s}, & d \le 0.5\text{~nm} \text{, average over the pore `surface' } 
  \end{cases}
\end{equation}
and
\begin{align}
  \beta_{\text{p}} &=
  \begin{cases}
    1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \\
    0, & \text{otherwise}
  \end{cases} \\
  \beta_{\text{b}} &=
  \begin{cases}
  1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \text{ and } d > 0.5 \\
  0, & \text{otherwise}
  \end{cases} \\
  \beta_{\text{s}} &=
  \begin{cases}
  1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \text{ and } d \le 0.5 \\
  0, & \text{otherwise}
  \end{cases}
\end{align}
with $d$ the distance from the nanopore wall and $r_\text{p}(z)$ is the radius of the pore at height $z$.

\newpage
\subsection{Weak forms of the ePNP-NS equations}

\paragraph{Poisson equation.}
The global potential distribution is described by the Poisson equation\cite{Lu-2012}
\begin{align} 
\label{eq:poisson}
\nabla \cdot \left( \absperm \relperm \nabla \potential \right) =
-\left( \scdpore + \scdion \right) \text{ ,}
\end{align}
with $\potential$ the electric potential, $\absperm$ the vacuum permittivity 
($8.85419\times10^{-12}\text{~F\,m}^{-1}$), and $\relperm$ the local relative permittivity of the medium. 
$\scdpore$ and $\scdion$ are the fixed (due to the pore) and mobile (due to the ions) charge distributions, 
respectively. To solve this equation with the finite element method, we must be derive its weak form by 
multiplication with an arbitrary test function and integration over the entire computational domain $\Omega$. 
For \cref{eq:poisson} with the potential test function $\psi$ this yields 
\begin{align}
\displaystyle\int_{\Omega} 
\left[
  \nabla \cdot \left( \absperm \relperm \nabla \potential \right)
\right]
\psi \,d\Omega
={}&
-
\displaystyle\int_{\Omega} \left[ \scdpore + \scdion \right] \psi \,d\Omega \notag \\
% newline
\displaystyle\int_{\Omega}
\left[
  \nabla \psi \cdot \left( \absperm \relperm \nabla \potential \right)
\right]
\,d\Omega
-
\displaystyle\int_{\Gamma}
\left[
  \psi \left( \absperm \relperm \nabla \potential \right) \cdot \vec{n}
\right]
\,d\Gamma
={}&
\displaystyle\int_{\Omega} \left[ \psi \scdpore \right] \,d\Omega
+
\displaystyle\int_{\Omega} \left[ \psi \scdion \right] \,d\Omega
\text{ ,}
\end{align}
with $\Gamma$ the boundary of $\Omega$ and $\vec{n}$ the normal vector on $\Gamma$.

\paragraph{Size-modified Nernst-Planck equation.}
The total ionic flux $\flux_{i}$ of ion $i$ expressed by the size-modified Nernst-Planck equation 
(smNPE)\cite{Lu-2012} includes a diffusive, electromigratory, steric and convective component
\begin{equation}
\label{eq:smpnp}
\nabla\cdot\flux_{i} = \nabla \cdot 
\left(
  - \diffusion_{i}\nabla\concentration_{i}
  - \chargen_{i}\mobility_{i}\concentration_{i}\nabla\potential
  - \beta_{i} \concentration_{i} 
  + \velocity\concentration_{i}
  \right) = 0  \text{\quad with }
  \beta_{i} = \frac
          {\ionsize_{i}^3/\ionsize_{0}^3 \displaystyle\sum_{j} \ionsize_{j}^3 \nabla \concentration_{j}}
          {1 - \displaystyle\sum_{j} \avogadro \ionsize_{j}^3 \concentration_{j}}
  \text{ ,}
\end{equation}
with ion diffusion coefficient $\diffusion_{i}$, concentration $\concentration_{i}$, charge number 
$\chargen_{i}$, mobility $\mobility_{i}$, electrostatic potential $\potential$, steric saturation factor 
$\beta_{i}$ and fluid velocity $\velocity$. $\avogadro$ is Avogadro's constant 
($6.022\times10^{23}\text{~mol}^{-1}$) and $\ionsize_{i}$ and $\ionsize_{0}$ are the limiting cubic diameters 
for ions and water, respectively. Using the ion concentration test function $d_i$, the weak form of 
\cref{eq:smpnp} becomes 
expressed as
\begin{align}
% raw weak expression
\displaystyle\int_{\Omega} \left[ \nabla \cdot \flux_{i} \right] d_{i} \,d\Omega ={}&
% apply Gauss theorem
\displaystyle\int_{\Omega} \left[\nabla d_{i} \cdot \flux_{i} \right]\,d\Omega 
+
\displaystyle\int_{\Gamma} \left[ d_{i} \flux_{i} \cdot \vec{n} \right]\,d\Gamma \\
={}&
-
\displaystyle\int_{\Omega}
\left[
  \nabla d_{i} \cdot \diffusion_{i}\nabla\concentration_{i}
\right]\,d\Omega
-
\displaystyle\int_{\Omega}
\left[
  \nabla d_{i} \cdot \chargen_{i}\mobility_{i}\concentration_{i}\nabla\potential
\right]\,d\Omega \notag \\
& -
\displaystyle\int_{\Omega}
\left[
  \nabla d_{i} \cdot \beta_{i} \concentration_{i}
\right]\,d\Omega
+
\displaystyle\int_{\Omega}
\left[
  \nabla d_{i} \cdot \velocity \concentration_{i}
\right]\,d\Omega \notag \\
& +
\displaystyle\int_{\Gamma}
\left[
  d_{i} \diffusion_{i} \nabla \concentration_{i} \cdot \vec{n}
\right]\,d\Gamma
+
\displaystyle\int_{\Gamma}
\left[
  d_{i} \chargen_{i} \mobility_{i} \concentration_{i} \nabla \potential \cdot \vec{n}
\right]\,d\Gamma \notag \\
& +
\displaystyle\int_{\Gamma}
\left[
  d_{i} \beta_{i} \concentration_{i}
  \cdot \vec{n}
\right]\,d\Gamma
+
\displaystyle\int_{\Gamma}
\left[
  d_{i} \velocity \concentration_{i} \cdot \vec{n}
\right]\,d\Gamma \notag \\
={}& 0 \notag
\text{ .}
\end{align}

\paragraph{Variable density and viscosity Navier-Stokes equation.}
The laminar fluid flow of an incompressible fluid with a variable density and viscosity is given 
by the system of equations\cite{Axelsson-2015}
\begin{align}
\label{eq:ns_density_continuity}
\vec{\vel} \cdot \nabla \density ={}& 0 \\
\label{eq:ns_conservation}
\left( \vec{\vel} \cdot \nabla \right) \left( \density\vec{\vel} \right) 
+ \nabla \cdot \hydrostresstensor ={}& \vec{\force} \text{\quad with }
\hydrostresstensor =
  \pressure\identity - \viscosity\left[\nabla\vec{\vel}+\left(\nabla\vec{\vel} \right)^\mathsf{T}\right] 
\\
\label{eq:ns_velocity_continuity}
\nabla \cdot \left( \density\vec{\vel} \right) - \vec{\vel} \cdot \nabla \density &= 0
\end{align}
with fluid velocity $\vec{\vel}$, density $\density$, viscosity $\viscosity$, pressure $\pressure$, body 
force $\vec{\force}$ and hydrodynamic stress tensor $\hydrostresstensor$.
The pressure test function $q$ is used to derive the weak forms of \cref{eq:ns_density_continuity}
\begin{align}
% Density continuity
\displaystyle\int_{\Omega} \left[ \vec{\vel} \cdot \nabla \density \right] q \,d\Omega ={}&
\displaystyle\int_{\Omega} \left[ q \vec{\vel} \cdot \nabla \density \right] \,d\Omega = 0 \text{ ,}
\end{align}
and \cref{eq:ns_velocity_continuity}
\begin{align}
% Velocity continuity
\displaystyle\int_{\Omega}
\left[\nabla \cdot \left( \density \vec{\vel} \right) - \vec{\vel} \cdot \nabla \density \right] q
\,d\Omega ={}&
\displaystyle\int_{\Omega}
\left[\nabla \cdot \left( \density \vec{\vel} \right) \right] q \,d\Omega
-
\displaystyle\int_{\Omega}
\left[ q \vec{\vel} \cdot \nabla \density \right] \,d\Omega \\
={}&
\displaystyle\int_{\Omega}
\left[\nabla q \cdot \left( \density \vec{\vel} \right) \right] \,d\Omega
-
\displaystyle\int_{\Gamma}
\left[ q \left( \density \vec{\vel} \right) \cdot \vec{n} \right] \,d\Gamma
-
\displaystyle\int_{\Omega}
\left[ q \vec{\vel} \cdot \nabla \density \right] \,d\Omega \notag
\text{ .}
\end{align}
For \cref{eq:ns_conservation} we use the velocity test function $\vec{v}=\left[v_r, v_\phi, v_z\right]$
\begin{align}
\displaystyle\int_{\Omega}
\left[
  \left( \vec{\vel} \cdot \nabla \right) \left( \density\vec{\vel} \right) + \nabla \cdot \hydrostresstensor
\right]
\cdot \vec{v} \,d\Omega
={}&
\displaystyle\int_{\Omega} \vec{\force} \cdot \vec{v} \,d\Omega \\
% Conservation
\displaystyle\int_{\Omega}
\left[
  \left( \vec{\vel} \cdot \nabla \right) \left( \density\vec{\vel} \right) \cdot\vec{v}
\right]
\,d\Omega
+
\displaystyle\int_{\Omega}
\left[
\hydrostresstensor \cdot \nabla\vec{v}
\right]
\,d\Omega
+
\displaystyle\int_{\Gamma}
\left[
\vec{v} \cdot
\hydrostresstensor
\cdot\vec{n}
\right]
\,d\Gamma
={}& \displaystyle\int_{\Omega} \vec{\force} \cdot \vec{v} \,d\Omega \notag
\text{ .}
\end{align}


\newpage
\section{Extented results}
\subsection{Voltage dependency of the conductance log-log slopes}
The values of the fitted power-law exponent $\gamma$ for both the simulated (ePNP-NS) and experimental data 
in the low and high concentration regimes can be found in \cref{fig:conductance_loglogslopes}.
\input{fig_conductance_loglogslopes.tex}

\subsection{Peak values of the radial potential profiles inside ClyA}
The peak values of the radial electrostatic potential $\radpot$ at the \cis\ entry, middle of the lumen and 
the \trans\ constriction for $0.005$, $0.05$, $0.15$, $0.5$ and $5$~M \ce{NaCl} are summarized in 
\cref{tab:radial_potential}.
\input{tab_radial_potential.tex}

\subsection{Pressure distribution inside ClyA}
The electro-osmotic pressure distribution inside ClyA, as a consequence of the strong local enhancement of 
the ion concentration, is given in \cref{fig:pressure}.
\input{fig_pressure.tex}


%-------------------------------------------------------------------------------
% SUPPORTING INFORMATION
%-------------------------------------------------------------------------------
\bibliography{modeling1}

\end{document}
