%-------------------------------------------------------------------------------
% PREAMBLE AND DOCUMENT FORMATTING
%-------------------------------------------------------------------------------
\documentclass[journal=ancac3, manuscript=suppinfo, etalmode=truncate,maxauthors=0]{achemso}
\setkeys{acs}{etalmode=truncate,maxauthors=0}

% PACKAGES
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{textgreek}

\usepackage[version=4]{mhchem}
\usepackage[alsoload=synchem,separate-uncertainty=true,multi-part-units=single,tight-spacing=true]{siunitx}

\usepackage{float}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage{array, booktabs, tabularx, multirow}

% CAPTION FORMATTING
\usepackage[font=footnotesize,labelfont=bf,labelsep=period]{caption} 							% format single-image captions and table titles
\captionsetup[table]{singlelinecheck=false,font=footnotesize,labelfont=bf}
\usepackage[font=footnotesize,labelfont=bf,labelsep=period]{subcaption} 						% format subfigure captions
%\DeclareCaptionSubType*[alph]{figure}
%\renewcommand\thesubfigure{\thefigure\alph{subfigure}}
\captionsetup[subfigure]{labelfont=bf,textfont=normalfont,labelformat=simple,singlelinecheck=false}

% CROSS-REFERENCE FORMATTING
% For use with the cleveref package
% Define the format of Figure, Table, Equation, and Section cross-references in the text
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{cleveref}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Tab.}{Tabs.}
\Crefname{table}{Table}{Tables}
\crefname{equation}{Eq.}{Eqs.}
\Crefname{equation}{Equation}{Equations}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}


% REFERENCES
\renewcommand*{\bibfont}{\normalfont\small}


%-------------------------------------------------------------------------------
% CUSTOM COMMANDS
%-------------------------------------------------------------------------------


\input{../shared/additional_commands.tex}


%-------------------------------------------------------------------------------
% TITLE AND AUTHOR LIST
%-------------------------------------------------------------------------------
\title{Modeling of Ion and Water Transport in the Biological Nanopore ClyA}
\input{../shared/authors.tex}


%===============================================================================
%
% MAIN DOCUMENT BEGINS
%
%===============================================================================
\begin{document}

\maketitle

\newpage
\section{Extended materials and methods}

\subsection{Fitting of the electrolyte properties}
\input{fig_corrections.tex}
The parameters of the fitting functions used to interpolate the experimental ion
diffusion coefficients, mobilities and transport numbers, and the electrolyte
viscosity, density and relative permittivity are given in
\cref{tab:corrections_parameters} and the resulting curves are plotted in
\cref{fig:corrections}. Note that since most of these functions are merely
empirical fits with no physical meaning, and that they are solely used to
interpolate and represent the experimental data.

Finally, for the concentration dependence of the relative permittivity we made
use of the model proposed by Gavish et al.\cite{Gavish-2016}
%
\begin{align}
\relperm(\dconc) = \permittivity_{r,0} - \left(\permittivity_{r,0} - \permittivity_{r,ms}\right) L \left( \dfrac{3\alpha}{\permittivity_{r,0} - \permittivity_{r,ms}} \dconc \right)
\end{align}
%
\input{tab_corrections_parameters.tex}

\newpage
\subsection{Surface integration to compute pore averaged values}

The average pore values for quantity of interest $X$ was computed by
%
\begin{align}
  \left< X \right>_{\alpha} =
    \displaystyle\frac{\displaystyle\iint_{V_{\alpha}} \beta_{\alpha} X \,dr\,dz}
                      {\displaystyle\iint_{V_{\alpha}} \beta_{\alpha} \,dr\,dz}
  \text{ ,}
\end{align}
%
where
%
\begin{equation}
  \alpha=
  \begin{cases}
    \text{p}, & d \ge 0  \text{~nm} \text{, average over the entire pore} \\
    \text{b}, & d > 0.5  \text{~nm} \text{, average over the pore `bulk' }  \\
    \text{s}, & d \le 0.5\text{~nm} \text{, average over the pore `surface' }
  \end{cases}
\end{equation}
%
and
%
\begin{align}
  \beta_{\text{p}} &=
  \begin{cases}
    1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \\
    0, & \text{otherwise}
  \end{cases} \\
  \beta_{\text{b}} &=
  \begin{cases}
  1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \text{ and } d > 0.5 \\
  0, & \text{otherwise}
  \end{cases} \\
  \beta_{\text{s}} &=
  \begin{cases}
  1, & \text{if}\ -1.85\le z \le 12.25  \text{ and } r \le r_\text{p}(z) \text{ and } d \le 0.5 \\
  0, & \text{otherwise}
  \end{cases}
\end{align}
%
with $d$ the distance from the nanopore wall and $r_\text{p}(z)$ is the radius of the pore at height $z$.


\newpage
\subsection{Weak forms of the ePNP-NS equations}
To solve partial differential equations with the finite element method, we must
be derive their weak form. This is achieved through multiplication of the
equation with an arbitrary test function and integration over their relevant
domains and boundaries (\cref{fig:domains_an_boundaries}). The full
computational domain of our model ($\Omega$) is subdivided into domains for the
pore ($\Omega_p$), the lipid bilayer ($\Omega_m$) and the electrolyte reservoir
($\Omega_w$). The relevant boundaries of these domains are also indicated, i.e.
the reservoir's exterior edges at the \cis\ $\Gamma_{w,c}$) and \trans\
$\Gamma_{w,t}$) sides, the outer edge of the lipid bilayer ($\Gamma_{m}$) and
the interface of the fluid with the nanopore and the bilayer ($\Gamma_{p+m}$).

The following paragraphs detail the equations that were used in this paper, together with their corresponding
weak forms.

\input{fig_domains_and_boundaries.tex}

\paragraph{Poisson equation.}
%
The global potential distribution is described by the Poisson equation
(PE)\cite{Lu-2012}
%
\begin{align}
\label{eq:poisson}
\nabla \cdot \displacement = -\left( \scdpore + \scdion \right)
\text{\quad with }
\displacement = \absperm \relperm \nabla \potential
\text{ ,}
\end{align}
%
with $\displacement$ the electrical displacement field, $\potential$ the
electric potential, $\absperm$ the vacuum permittivity
($8.85419\times10^{-12}\text{~F\,m}^{-1}$), and $\relperm$ the local relative
permittivity of the medium. $\scdpore$ and $\scdion$ are the fixed (due to the
pore) and mobile (due to the ions) charge distributions, respectively.

Multiplication of \cref{eq:poisson} with the potential test function $\psi$ and
integration over the entire model $\Omega=\Omega_w+\Omega_p+\Omega_m$ gives
%
\begin{align}
\displaystyle\int_{\Omega}
\left[
  \nabla \cdot \displacement
\right]
\psi \,d\Omega
={}&
- \displaystyle\int_{\Omega} \left[ \scdpore + \scdion \right] \psi \,d\Omega \text{ ,}
\end{align}
%
which, after applying the Gauss divergence theorem, yields the final weak
formulation
%
\begin{align}
\displaystyle\int_{\Omega}
\left[
  \nabla \psi \cdot \displacement
\right]
\,d\Omega
- \displaystyle\int_{\Gamma_\text{PE}}
\left[
  \psi \displacement \cdot \vec{n}
\right]
\,d\Gamma_\text{PE}
={}&
\displaystyle\int_{\Omega} \left[ \psi \scdpore \right] \,d\Omega
+
\displaystyle\int_{\Omega} \left[ \psi \scdion \right] \,d\Omega
\text{ ,}
\end{align}
%
with boundaries $\Gamma_\text{PE}=\Gamma_{w,c}+\Gamma_{w,t}+\Gamma_m$ and
$\vec{n}$ their normal vector. The boundary integrals at $\Gamma_{w,c}$ and
$\Gamma_{w,t}$ are evaluated using the Dirichlet boundary conditions (BCs)
$\potential=0$ and $\potential=\vbias$, respectively. A zero charge BC, $\vec{n}
\cdot \displacement = 0$, is used for the integral at $\Gamma_{m}$.

\paragraph{Size-modified Nernst-Planck equation.}
%
The total ionic flux $\flux_{i}$ of ion $i$ at steady-state is expressed by the
size-modified Nernst-Planck equation (smNPE)\cite{Lu-2012}
%
\begin{equation}
\label{eq:smnp}
\pd{\concentration_{i}}{t} = - \nabla\cdot\flux_{i} = - \nabla \cdot
\left(
  \diffusion_{i}\nabla\concentration_{i}
  + \chargen_{i}\mobility_{i}\concentration_{i}\nabla\potential
  + \vec{\beta_{i}} \concentration_{i}
  - \velocity\concentration_{i}
\right)
\text{\quad with }
\vec{\beta_{i}} =
  \frac{\ionsize_{i}^3/\ionsize_{0}^3 \displaystyle\sum_{j} \ionsize_{j}^3 \nabla \concentration_{j}}
      {1 - \displaystyle\sum_{j} \avogadro \ionsize_{j}^3 \concentration_{j}}
\text{ ,}
\end{equation}
%
with ion diffusion coefficient $\diffusion_{i}$, concentration
$\concentration_{i}$, charge number $\chargen_{i}$, mobility $\mobility_{i}$,
electrostatic potential $\potential$, steric saturation factor $\beta_{i}$ and
fluid velocity $\velocity$. $\avogadro$ is Avogadro's constant
($6.022\times10^{23}\text{~mol}^{-1}$) and $\ionsize_{i}$ and $\ionsize_{0}$ are
the limiting cubic diameters for ions and water, respectively. Using the ion
concentration test function $d_i$, the weak form of \cref{eq:smnp} becomes
%
\begin{align}
% Raw integral
\displaystyle\int_{\Omega_w} \left[ \pd{\concentration_{i}}{t} \right] d_{i} \,d\Omega_w ={}&
\displaystyle\int_{\Omega_w} \left[ - \nabla \cdot \flux_{i} \right] d_{i} \,d\Omega_w \notag \\
% Gauss divergence theorem
\displaystyle\int_{\Omega_w} \left[ d_{i} \pd{\concentration_{i}}{t} \right] \,d\Omega_w ={}&
\displaystyle\int_{\Omega_w} \left[\nabla d_{i} \cdot \flux_{i} \right]\,d\Omega_w
- \displaystyle\int_{\Gamma_\text{NP}}
\left[ d_{i} \flux_{i} \cdot \vec{n} \right]\,d\Gamma_\text{NP} \notag \\
% Fill in fluc components
={}&
% Domain
\displaystyle\int_{\Omega_w}
\left[
  \nabla d_{i} \cdot
  \left(
    \diffusion_{i}\nabla\concentration_{i}
    + \chargen_{i}\mobility_{i}\concentration_{i}\nabla\potential
    + \vec{\beta_{i}} \concentration_{i}
    - \velocity \concentration_{i}
  \right)
\right]
\,d\Omega_w \notag \\
% Boundary
& - \displaystyle\int_{\Gamma_\text{NP}}
\left[
  d_{i}
  \left(
    \diffusion_{i} \nabla \concentration_{i}
    + \chargen_{i} \mobility_{i} \concentration_{i} \nabla \potential
    + \vec{\beta_{i}} \concentration_{i}
    - \velocity \concentration_{i}
  \right)
  \cdot \vec{n}
\right]
\,d\Gamma_\text{NP}
\text{ .}
\end{align}
%
The integrals on the boundaries $\Gamma_\text{NP} =
\Gamma_{w,c}+\Gamma_{w,t}+\Gamma_{p+m}$ are evaluated using the Dirichlet
boundary condition $\concentration_{i} = \cbulk$ for $\Gamma_{w,c}$ and
$\Gamma_{w,t}$, and the no flux boundary condition $\vec{n} \cdot \flux_{i} = 0$
for $\Gamma_{p+m}$

\paragraph{Variable density and viscosity Navier-Stokes equation.}
%
The steady-state, laminar fluid flow of an incompressible fluid with a variable
density and viscosity is given by the system of equations\cite{Axelsson-2015}
%
\begin{align}
\label{eq:ns_density_continuity}
\velocity \cdot \nabla \density ={}& 0 \\
\label{eq:ns_conservation}
\left( \velocity \cdot \nabla \right) \left( \density\velocity \right)
+ \nabla \cdot \hydrostresstensor ={}& \vec{\force} \text{\quad with }
\hydrostresstensor =
  \pressure\identity - \viscosity\left[\nabla\velocity+\left(\nabla\velocity \right)^\mathsf{T}\right]
\\
\label{eq:ns_velocity_continuity}
\nabla \cdot \left( \density\velocity \right) - \velocity \cdot \nabla \density ={}& 0
\text { ,}
\end{align}
%
with fluid velocity $\velocity$, density $\density$, hydrodynamic stress tensor
$\hydrostresstensor$, viscosity $\viscosity$, pressure $\pressure$ and body
force $\vec{\force}$. The pressure test function $q$ is used to derive the weak
forms of \cref{eq:ns_density_continuity}
%
\begin{align}
% Density continuity
\displaystyle\int_{\Omega_w} \left[ \velocity \cdot \nabla \density \right] q \,d\Omega_w ={}&
\displaystyle\int_{\Omega_w} \left[ q \velocity \cdot \nabla \density \right] \,d\Omega_w = 0 \text{ ,}
\end{align}
and \cref{eq:ns_velocity_continuity}
\begin{align}
% Velocity continuity
\displaystyle\int_{\Omega_w}
\left[\nabla \cdot \left( \density \velocity \right) - \velocity \cdot \nabla \density \right] q
\,d\Omega_w ={}&
\displaystyle\int_{\Omega_w}
\left[\nabla \cdot \left( \density \velocity \right) \right] q \,d\Omega_w
-
\displaystyle\int_{\Omega_w}
\left[ q \velocity \cdot \nabla \density \right] \,d\Omega_w \\
={}&
\displaystyle\int_{\Omega_w}
\left[\nabla q \cdot \left( \density \velocity \right) \right] \,d\Omega_w
-
\displaystyle\int_{\Gamma_\text{NS}}
\left[ q \left( \density \velocity \right) \cdot \vec{n} \right] \,d\Gamma_\text{NS}
-
\displaystyle\int_{\Omega_w}
\left[ q \velocity \cdot \nabla \density \right] \,d\Omega_w \notag
\text{ ,}
\end{align}
%
while for \cref{eq:ns_conservation} we use the velocity test function
$\vec{v}=\left[v_r, v_\phi, v_z\right]$
%
\begin{align}
\displaystyle\int_{\Omega_w}
\left[
  \left( \velocity \cdot \nabla \right) \left( \density\velocity \right) + \nabla \cdot \hydrostresstensor
\right]
\cdot \vec{v} \,d\Omega_w
={}&
\displaystyle\int_{\Omega_w} \vec{\force} \cdot \vec{v} \,d\Omega_w \\
% Conservation
\displaystyle\int_{\Omega_w}
\left[
  \left( \velocity \cdot \nabla \right) \left( \density\velocity \right) \cdot\vec{v}
\right]
\,d\Omega_w
-
\displaystyle\int_{\Omega_w}
\left[
\hydrostresstensor \cdot \nabla\vec{v}
\right]
\,d\Omega_w
+
\displaystyle\int_{\Gamma_\text{NS}}
\left[
\vec{v} \cdot
\hydrostresstensor
\cdot\vec{n}
\right]
\,d\Gamma_\text{NS}
={}& \displaystyle\int_{\Omega_w} \vec{\force} \cdot \vec{v} \,d\Omega_w \notag
\text{ ,}
\end{align}
%
with boundaries $\Gamma_\text{NS} = \Gamma_{w,c}+\Gamma_{w,t}+\Gamma_{p+m}$. The
no-slip Dirichlet BC $\velocity = 0$ is applied to $\Gamma_{p+m}$, and the no
normal stress $\hydrostresstensor\vec{n}=0$ is used for $\Gamma_{w,c}$ and
$\Gamma_{w,t}$.

\newpage
\section{Extended results}

\subsection{Peak values of the radial potential profiles inside ClyA}

The peak values of the radial electrostatic potential $\radpot$ at the \cis\
entry, middle of the lumen and the \trans\ constriction for $0.005$, $0.05$,
$0.15$, $0.5$ and $5$~M \ce{NaCl} are summarized in \cref{tab:radial_potential}.
\input{tab_radial_potential.tex}

\subsection{Pressure distribution inside ClyA}
The electro-osmotic pressure distribution inside ClyA, as a consequence of the
strong local enhancement of the ion concentration, is given in
\cref{fig:pressure}.
\input{fig_pressure.tex}


%-------------------------------------------------------------------------------
% SUPPORTING INFORMATION
%-------------------------------------------------------------------------------
\bibliography{../shared/bibliography}

\end{document}
